---
title: "Fort Worth Climate Change"
description: |
  Fort Worth climate change based on precipitation, humidity, and temperature from 1981-2022
author:
  - name: Ethan Campbell
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: TRUE
    theme: bootstrap.min.css
---


      

<style>
html {
  scroll-behavior: smooth;
}
#TOC {
  position: relative;
  z-index: 50;
  background: #ebebeb;
  padding: 10px;
  border-radius: 5px;
}
@media screen and (max-width: 900px) {
#TOC {
    position: relative;
  }
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plyr)
library(dplyr)
library(rsconnect)
library(tidyr)
library(ggridges)
library(viridis)
library(viridisLite)
library(kableExtra)
library(gganimate)
library(distill)
library(lubridate)
library(leaflet)
library(tibble)
library(ggplot2)
library(ggrepel)
library(plotly)
library(ggthemes)
```


# Introduction
  Climate change has been a reoccurring question for as long as I can remember and recently the discussion of climate change in Texas has grown exponentially. Due to recent events such as the polar vortex people are considering the effects of climate change in a new light. With this data set I would like to analyze the changes in climate over the last 40 years to gain more insight into the question of climate change in Texas.

# Data

  The data that I used in this study was gathered from NASA prediction of Worldwide Energy Resources. I extracted the data set from the web mapping application. This data set provided the information of wind, temperature, surface pressure, precipitation, and humidity. The factors that I will be looking at are temperature, humidity, and precipitation as these relate to the climate change and I can also look at how they correlate with one another.I have included an interactive map of Fort Worth Tx for those unfamiliar with the area. Further down the road I would like to implement the other factors for a full comparison and set up a system that will update with each passing day. In the first data table below are the data types of each column within this study.

```{r}
leaflet() %>%
  addTiles() %>%
  addMarkers(lng=-97.3225, lat=32.756, popup="Fort Worth")

```



# Manipulating data
  I started by inputting the data and viewing it to identify any changes that needed to be made and noticed it needed some cleaning before the analytic process could begin. In this study we will focus on three variables which are temperature, humidity, and precipitation. Before I could start using the data there needed to be some changes. The first change was creating a date column using lubridate. Afterwards I created an annual column for each variable by utilizing the mean spread out over a one year period for each year. For the visualization process, I added in a months abbreviation which instead of the numeric number for a month it displays the abbreviation. 

```{r}

Fort_Worth_2022 <- read.csv("Fort_Worth_climate_with_day.csv", skip = 14)
Fort_Worth <- read.csv("Fort_Worth_climate.csv", skip = 18)


Month_combined <- Fort_Worth %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_AVG",
)

Para_split <- Month_combined %>%
  pivot_wider(names_from = PARAMETER,
              values_from = Month_AVG,
              )


FW_Updated <- Fort_Worth_2022 %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_Speed = WS2M) %>%
  dplyr::rename(Surface_Pressure = PS)


FW_Updated$Date <- with(FW_Updated, ymd(sprintf('%04d%02d%02d', YEAR, MO, Day)))


YEAR <- format(as.Date(FW_Updated$Date), format = "%Y")
 
Means_variables <- ddply(FW_Updated, .(YEAR), summarise,
      Annual_Temperature = mean(Temperature),
      Annual_Humidity = mean(Humidity),
      Annual_Precipitation = mean(PRECTOTCORR))

Final_FW <- merge(FW_Updated, Means_variables, by = "YEAR")


Final_FW <- transform(Final_FW, MonthAbb = month.abb[MO])

rmarkdown::paged_table(Final_FW)



```


# Temperature

  Starting off we mutated to adjust from Celsius to Fahrenheit and selected the important columns that would impact this section. I decided to start by seeing when temperature was the highest and lowest and found that it was at its peak in August of 2011 and its lowest in December of 1989. The graphs show the change in temperature by month with the first showing amount of occurrences for each month being over 5 degrees. The second shows the full range of temperature per month. We notice a really large standard deviation indicating that there is a wide range in values which is to be expected considering the annual range of temperature is rather large. However, I did not expect it to be 17.3 which was an interesting conclusion. 

```{r}
Temperature_Final <- Final_FW %>%
  mutate(Temperature = Temperature * 9/5 + 32) %>%
  mutate(Annual_Temp = Annual_Temperature * 9/5 + 32) %>%
  select(Date, YEAR, MO, Day, MonthAbb, Temperature, Annual_Temp) %>%
  na.omit(Temperature)


# When was Temperature the highest (Aug of 2011)

TF_highest <- Temperature_Final %>%
  select(Date,Temperature, Annual_Temp) %>%
  arrange(desc(Temperature)) %>%
  slice(1:12)

kable(TF_highest, digits = 4, align = "ccccccc", col.names = c("Date", "Temperature", "Annual Temperature"), caption = "Highest Temperature Since 1981") %>%
  kable_styling(font_size = 16) %>%
  row_spec(c(1,1,1), background = "red")

# When was Temperature the lowest (Dec of 1989)

TF_lowest <- Temperature_Final %>%
  select(Date, Temperature, Annual_Temp) %>%
  arrange(Temperature) %>%
  na.omit(Temperature) %>%
  slice(1:12)

kable(TF_lowest, digits = 4, align = "ccccccc", col.names = c("Date", "Temperature", "Annual Temperature"), caption = "Lowest Temperature Since 1981") %>%
  kable_styling(font_size = 16) %>%
  row_spec(c(1,1,1), background = "cadetblue")

TF_stats <- Temperature_Final %>%
  select(Date, Temperature, Annual_Temp) %>%
  mutate(Mean = mean(Temperature)) %>%
  mutate(Standard_Deviation = sd(Temperature)) %>%
  mutate(Median = median(Temperature)) %>%
  slice(1:12)

kable(TF_stats, digits = 4, align = "ccccccc", col.names = c("Date", "Temperature", "Annual Temperature", "Mean", "Standard Deviation", "Median"), caption = "Statistical Temperature Data") %>%
  kable_styling(font_size = 16) %>%
  row_spec(c(1,1,1,1,1,1))

Temperature_Final %>%
drop_na(Temperature) %>%
filter(Temperature > 0) %>%
ggplot(aes(Temperature, fill = MonthAbb), label = "Month") +
geom_density(mapping = aes(color = Temperature), alpha = 0.5) +
facet_wrap(~MonthAbb) +
theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Percent Changed') + xlab('Years') +
  labs(title = "Temperature Density Graph Greater Than 0 [F]", caption = "Facet Grid of each month exceeding the temperature of 0 [F]")



Temperature_Final %>%
ggplot(mapping = aes(x = Temperature, y = MO, group = MO, fill = ..x..)) +
geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01,
alpha = 5) +
scale_fill_viridis(name = "Temp. [F]", option = "C") +
theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Months') + xlab('Temperature [F]') +
  labs(title = "Mean Temperature Range for Each Month", caption = "")


```

## Temperature percent change
  Here we broke the temperature column down to analyze the percent change from year to year. We noticed that there was a fairly even back and forth in terms of percent change every 3 years. The next portion is the Temperature change difference from 1981. Here we see that the temperature actually increased by almost 2%. To make the visualization portion easier to the eye I made it every 3 years instead of 1. The temperature since 1981 tends to lean towards a higher number more frequently than a lower number however, the amount isn't significant enough to call it a complete temperature change and can be chalked up to a natural variation in temperature.

```{r}
TF <- Temperature_Final %>%
  distinct(YEAR, Annual_Temp)

TF <- TF %>%
  slice(which(row_number() %% 3 == 1))

YearOneprepTemp <- TF[1,c("Annual_Temp")]

Temp_change <- TF %>%
  dplyr::mutate(Previous = lag(Annual_Temp),
                Next_temp = lead(Annual_Temp),
                change_temp = Annual_Temp - Previous,
                Percent_temp = (change_temp/Previous)* 100, 
                Percent_change_temp = (change_temp/lag(Annual_Temp) -1) * 100,
                TChange_from_year_one = (Annual_Temp/YearOneprepTemp - 1) * 100)


Temp_change <- Temp_change %>%
  select(YEAR, Annual_Temp, Percent_temp, TChange_from_year_one)


ggplot(data = Temp_change) +
  geom_col(aes(x = YEAR, y = Percent_temp, fill = Percent_temp > 0), alpha = .2) +
  geom_text(aes(x = YEAR, y = Percent_temp, label = paste0(round(Percent_temp,2), "%")),size = 3, vjust = -.5) +
   theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Percent Changed') + xlab('Years') +
  labs(title = "Temperature Percent Change Over 40 Years [%]", caption = "3 Year time gaps") 


ggplot(data = Temp_change) +
  geom_col(aes(x = YEAR, y = TChange_from_year_one, fill = TChange_from_year_one > 0), alpha = .2) +
  geom_text(aes(x = YEAR, y = TChange_from_year_one, label =  paste0(round(TChange_from_year_one,2), "%")),size = 3, vjust = -.5) +
  theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Percent Changed') + xlab('Years') +
  labs(title = "Temperature Percent Change From 1981 [%]", caption = "3 Year time gaps")
```

# Precipitation

  Precipitation is one that I suspect is vastly different as the weather in Texas has been abnormal as of late showing signs of a higher precipitation rate. Just like the temperature we pulled the highest and lowest precipitation months. With October of 1981 being the highest with 15.63 inches and January of 1986 being the lowest of .0098 inches. The animated plot includes the precipitation of months by year. This shows us which months usually have the highest and lowest and what their range is. The standard deviation in this section is 2.18 which is still high however, considering the abnormality of precipitation in the last few years I am surprised this number isn't higher. The reason I say this is because my town went through a drought roughly 5-7 years ago and then we went through a flood season and finally the major winter storm here. I expected the range between these events to create a larger standard deviation. 

```{r}
Precipitation_Final <- Para_split %>%
  mutate(Precipitation_annual = ANN / 25.4) %>%
  mutate(Precipitation_Monthly = PRECTOTCORR_SUM / 25.4) %>%
  select(YEAR, MONTH, Precipitation_Monthly, Precipitation_annual) %>%
  na.omit(Precipitation_Monthly)


# When was precipitation the highest (OCT of 1981)

PF_highest <- Precipitation_Final %>%
  select(YEAR, MONTH, Precipitation_Monthly, Precipitation_annual) %>%
  arrange(desc(Precipitation_Monthly)) %>%
  na.omit(Precipitation_Monthly) %>%
  slice(1:12)

kable(PF_highest, digits = 4, align = "ccccccc", col.names = c("Year", "Month","Precipiation", "Annual Precipitation"), caption = "Highest Monthly Mean Precipitation since 1981") %>%
  kable_styling(font_size = 16) %>%
  row_spec(c(1,1,1))
# When was it the lowest (Jan of 1986)

PF_lowest <- Precipitation_Final %>%
  select(YEAR, MONTH, Precipitation_Monthly, Precipitation_annual) %>%
  arrange(Precipitation_Monthly) %>%
  na.omit(Precipitation_Monthly) %>%
  slice(1:12)

kable(PF_lowest, digits = 4, align = "ccccccc", col.names = c("Year", "Month","Precipiation", "Annual Precipitation"), caption = "Lowest Monthly Mean Precipitation since 1981") %>%
  kable_styling(font_size = 16) %>%
  row_spec(c(1,1,1))


PF_stats <- Precipitation_Final %>%
  select(YEAR, MONTH, Precipitation_Monthly, Precipitation_annual) %>%
  mutate(Mean = mean(Precipitation_Monthly)) %>%
  mutate(Standard_Deviation = sd(Precipitation_Monthly)) %>%
  mutate(Median = median(Precipitation_Monthly)) %>%
  slice(1:12)

kable(PF_stats, digits = 4, align = "ccccccc", col.names = c("Year", "Month", "Monthly Precipitation", "Annual Precipitation", "Mean", "Standard Deviation", "Median"), caption = "Statistical Temperature Data") %>%
  kable_styling(font_size = 16) %>%
  row_spec(c(1,1,1,1,1,1,1))

Precip_animate <- ggplot(Precipitation_Final, aes(x = YEAR, y = Precipitation_Monthly, color = MONTH, group = MONTH)) +
    geom_path() +
    geom_point() +
    facet_wrap(~ MONTH) +
    theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Years') + xlab('Monthly Precipitation (Inches)') +
  labs(title = 'Precipitation Variation, Year: {frame_along}', caption = "Precipitation Change From 1981-2022") +
  theme(panel.spacing.x=unit(1.5, "lines"), panel.spacing.y = unit(1, "lines")) +
  transition_reveal(along = YEAR) +
    ease_aes('linear')


animate(Precip_animate, 100, 9)

warning()
```

## Precipitation precent change

  Here we find the percent change every 3 years and one thing we notice is that there has been an increase in precipitation in the last decade. For some reason the change from year 1 did not like either the code or the data type of my column because it kept switching the data type to data.table. I was unsure how to resolve this problem at first and added the interactive chart however, after some messing around I found a way to fix it. I had to remove the column name and incorporate the original value itself then divide by each year following. For some reason it didn't like the way I originally wrote it which was used for the other values however, we worked around it. Looking at this table we notice that there has been a decrease in precipitation by 13.48% which is very interesting. I expected this value to be positive .01 at the very lowest. However. when reading into the data we notice that 1981 had 42.6 inches of rain that year while 2020 only had 36.86 inches resulting in the negative number. My reasoning for the perception that there has been more precipitation is that we have gotten more precipitation in a shorter amount of time and less spanned out over the year. 

```{r}
PF <- Precipitation_Final %>%
  select(YEAR, Precipitation_annual)

PF <- PF %>%
  distinct(YEAR, Precipitation_annual)

PF <- PF %>%
  slice(which(row_number() %% 3 == 1))

Prep_Year_One_prep <- PF[1,c("Precipitation_annual")]

Pct_change <- PF %>%
  dplyr::mutate(Previous = lag(Precipitation_annual),
                Next = lead(Precipitation_annual),
                change = Precipitation_annual - Previous,
                Percent = (change/Previous)* 100, 
                Percent_change = (change/lag(Precipitation_annual) -1) * 100)

options(scipen = 999)

Pct_change$Change_from_year_one <- (PF$Precipitation_annual/42.60591 - 1) * 100

Pct_change <- Pct_change %>%
  select(YEAR, Precipitation_annual, Percent, Change_from_year_one)

ggplot(data = Pct_change) +
  geom_col(aes(x = YEAR, y = Percent, fill = Percent > 0), alpha = .2) +
  theme_classic() +
  geom_text(aes(x = YEAR, y = Percent, label = paste0(round(Percent,2), "%")),size = 3, vjust = -.5) + 
      theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Percent Changed') + xlab('Years') +
  labs(title = "Percipitation Percent Change Over 40 Years [%]", caption = "3 Year time gaps")


ggplot(data = Pct_change) +
 geom_col(aes(x = YEAR, y = Change_from_year_one, fill = Change_from_year_one > 0), alpha = .2) +
  theme_classic() +
  geom_text(aes(x = YEAR, y = Change_from_year_one, label = paste0(round(Change_from_year_one,2), "%")),size = 3, vjust = -.5) + 
      theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Percent Changed') + xlab('Years') +
  labs(title = "Percipitation Percent Change From 1981 [%]", caption = "3 Year time gaps")

Interactive_Precipitation_change <- Precipitation_Final %>%
  ggplot( aes(YEAR, Precipitation_annual, size = Precipitation_annual, color=Precipitation_annual)) +
  geom_point() +
    theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Annual Precipitation [Inches]') + xlab('Years') +
  labs(title = "Annual Precipitation Change Over 40 Years", caption = "")

ggplotly(Interactive_Precipitation_change)
```

# Humidity

  Out of all the variables humidity was the easiest to work with and was very straight forward as there were no major changes that needed to be made. Once more we looked at when humidity was the highest and lowest and noticed that in December of 2017 the humidity was it's highest of 97.44%. While in October of 2015 it was at it's lowest of 20%. The standard deviation reached 15.00 which was really high compared to the precipitation and I expected this number to be much lower. I created a bar graph to show which month reaches the highest average humidity and noticed that April was the leader. I also added a facet wrap split up by month to show each months change over the years. 

```{r, layout="1-body-outset"}
rmarkdown::paged_table(Humidity_Final <- Final_FW %>%
  select(Date, YEAR, MO, Day, MonthAbb, Humidity, Annual_Humidity) %>%
  na.omit(Humidity))

# When was humidity the highest (Jan of 1998)

HF_highest <- Humidity_Final %>%
  select(Date, Humidity, Annual_Humidity) %>%
  arrange(desc(Humidity)) %>%
  slice(1:12)

kable(HF_highest, digits = 4, align = "ccccccc", col.names = c("Date", "Humidity", "Annual Humidity"), caption = "Highest Humidity Percent In One Day Since 1981") %>%
  kable_styling(font_size = 16) %>%
  row_spec(c(1,1,1))

# When was humidity the lowest (Aug of 2000)

HF_lowest <- Humidity_Final %>%
  select(Date, Humidity, Annual_Humidity) %>%
  arrange(Humidity) %>%
  na.omit(Humidity) %>%
  slice(1:12)

kable(HF_lowest, digits = 4, align = "ccccccc", col.names = c("Date", "Humidity", "Annual Humidity"), caption = "Lowest Humidity Percent In One Day Since 1981") %>%
  kable_styling(font_size = 16) %>%
  row_spec(c(1,1,1))

HF_stats <- Humidity_Final %>%
  select(Date, Humidity, Annual_Humidity) %>%
  mutate(Mean = mean(Humidity)) %>%
  mutate(Standard_Deviation = sd(Humidity)) %>%
  mutate(Median = median(Humidity)) %>%
  slice(1:12)

kable(HF_stats, digits = 4, align = "ccccccc", col.names = c("Date", "Humdity", "Annual Humdity", "Mean", "Standard Deviation", "Median"), caption = "Statistical Humidity Data") %>%
  kable_styling(font_size = 16) %>%
  row_spec(c(1,1,1,1,1,1))


# change in humidity over years by each month

ggplot(data = Humidity_Final, mapping = aes(x = Humidity)) +
  geom_bar(mapping = aes(fill = MonthAbb), width = .5) +
    guides(
    color = guide_colorbar(
      nrow = 1,
      override.aes = list(size = 4)
    )
  ) +
  theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Occurances') + xlab('Humidity [%]') +
  labs(title = "Humidity Percent Occurance", caption = "")

# Facet wrap of the previous graph to separate them

ggplot(Humidity_Final, aes(YEAR, Humidity, colour = Humidity)) +
  geom_point(size = 0.5) +
  geom_smooth(mapping = aes(color = YEAR)) +
  theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 12)) + ylab('Humidity [%]') + xlab('Years') +
  labs(title = "Humidity Percent Change By Each Month", caption = "") +
  theme(panel.spacing.x=unit(.5, "lines") , panel.spacing.y=unit(.2,"lines")) +
  facet_wrap(~MonthAbb)
```

## Percent change 

  Once more Incorporated a percent change and did it every 3 years and here we noticed that over the last decade there has been an increase in humidity which correlates with the precipitation changes. However, when we look at the change from year 1 we notice that there has actually been a decrease of humidity by .03%. This was another interesting number because there is a correlation of 76.8%. With the major decrease in overall precipitation I expected this number to correlate with that change. 

```{r}

cor(Final_FW$Annual_Humidity,Final_FW$Annual_Precipitation)

HF <- Humidity_Final %>%
  select(YEAR, MO, Annual_Humidity, Humidity)

HF <- HF %>%
  distinct(YEAR, Annual_Humidity)
HF <- HF %>%
  slice(which(row_number() %% 3 == 1))
HYearOneprep <- HF[1,c("Annual_Humidity")]

HPct_change <- HF %>%
  dplyr::mutate(hPrevious = lag(Annual_Humidity),
                HNext = lead(Annual_Humidity),
                Hchange = Annual_Humidity - hPrevious,
                HPercent = (Hchange/hPrevious)* 100, 
                HPercent_change = (Hchange/lag(Annual_Humidity) -1) * 100,
                Hum_Change_from_year_one = (Annual_Humidity/HYearOneprep - 1) * 100)


HPct_change <- HPct_change %>%
  select(YEAR, Annual_Humidity, HPercent, Hum_Change_from_year_one)

ggplot(data = HPct_change) +
  geom_col(aes(x = YEAR, y = HPercent, fill = HPercent > 0), alpha = .2) +
  geom_text(aes(x = YEAR, y = HPercent, label = paste0(round(HPercent,2), "%")),size = 3, vjust = -.5) + 
    theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Percent Changed') + xlab('Years') +
  labs(title = "Humidity Percent Change From 1981 [%]", caption = "3 Year time gaps")


ggplot(data = HPct_change) +
  geom_col(aes(x = YEAR, y = Hum_Change_from_year_one, fill = Hum_Change_from_year_one > 0), alpha = .2) +
  geom_text(aes(x = YEAR, y = Hum_Change_from_year_one, label = paste0(round(Hum_Change_from_year_one,2), "%")),size = 3, vjust = -.5) + 
  theme_fivethirtyeight(base_size = 10, base_family = 'serif') +
  theme(axis.title = element_text(family = 'serif', size = 15)) + ylab('Percent Changed') + xlab('Years') +
  labs(title = "Humidity Percent Change Since 1981 [%]", caption = "3 Year time gaps")


```

# Conclusion: Has the climate changed?

  I want to start this conclusion with the definition of climate change to determine how this analysis conclusion will be dictated. “Climate change is a long-term change in the average weather patterns that have come to define Earth’s local, regional and global climates” (NASA). Fort Worth's current climate is humid subtropical which includes high temperatures, humid summers and mild to cool winters. With this in mind I will lay out the percent changes of each variable and discuss any major occurrences related to them. First temperature, when analyzing temperature we noticed a fairly consistent back and forth motion in yearly temperatures. Another point to note is comparing the year 1981 to 2022 we notice an increase of 1.78%. Due to the fact that this is not a consistent increase nor is the change significant we can assume this is within the range of the current climate. The next point is precipitation, when analyzing the changes over the years we noticed a little less back and forth but with larger numbers going in opposite directions. The difference from 1981 to 2022 was a decrease of 13.48% however, the precipitation amounts have been increasing by roughly 15% every 3 years. In about 2012 we had a decrease in temperature compared to the year 1981 of 47.12% and this negative amount has slowly been on the rise. We have two assumptions here, one 1981 was a high precipitation year or two there has simply been a decrease in the amount of precipitation in the last 40 years. One thing that makes me question option 1 is in 1990 we had an increase in precipitation of 9.61% which makes me lean more towards option two. I believe the precipitation has shown a climate in-between west Texas which is semi-arid and the typical humid subtropical of Fort Worth. The final section is humidity which has shown something similar to temperature which is a wave like motion of change over the years. The 1981 to 2022 change is -.03% which has been coming up from the -17.96% of 2012. The difference compared to 1981 has been mostly negative except for roughly 1990 again similar to precipitation. This is expect as there is roughly a 77% correlation rate between these two factors. Now with a change of -.03% we can chalk any changes of humidity in the last 40 years to the range variation of the humid subtropical climate. With all factors considered, I believe there has not been any major changes in climate and the difference we are seeing is simply the range inside of the humid subtropical climate. However, that is not to say that things could not change as we have seen with the precipitation. 

# Bibliography {.appendix} 

* NASA. Overview: Weather, Global Warming, and Climate Change. 2022. 17 April 2022. <https://climate.nasa.gov/resources/global-warming-vs-climate-change/>.

* R Core Team (2021). R: A language and environment for statistical
  computing. R Foundation for Statistical Computing, Vienna, Austria.
  URL <https://www.R-project.org/>.

* Research, NASA Langley. The POWER Project. 08 May 2021. 21 February 2022. <https://power.larc.nasa.gov.>.

  * (“These data were obtained from the NASA Langley Research Center (LaRC) POWER Project funded through the NASA Earth Science/Applied Science Program.”)

* Wickham, Hadley and Garrett Grolemund. R for Data Science. O'REILLY, 2017. <https://r4ds.had.co.nz>.




